Retrofitting Effect Handlers onto OCaml

### 2020-11-09 retro-concurrency.tex
Added web server results

@@

\section*{Evaluation Planning}

\subsection{Webserver benchmarks}

TODO:
\begin{itemize}
\item Aim: |R3| Do we have an efficient implementation of effects.
\item effects are a very nice way to program non-blocking IO for webservers
\item show that using effects doesn't impose undue overhead and that we are competitive with LWT, Async, Go with the wrk web benchmark
\end{itemize}

\subsection{Using fibers for task based parallelism}

TODO:
\begin{itemize}
\item Aim: |R3| Do we have an efficient implementation of effects.
\item minor alteration to domainslib to launch tasks in their own stack
\item effects are the right way to do some task base parallelism. Use the nqueens example and show the scaling we can achieve in parallel
\item show that using fibers doesn't have large detrimental impact to existing parallel applications: Take sandmark: minilight, nbody, LU decomposition
\end{itemize}

>>>>

\paragraph{Web server} As one of our primary goals of adding effect handlers to
OCaml is to support direct-style concurrency, we implement a full-fledged
HTTP/1.1 web server using effect handlers (|MC|). We use effect handlers to
implement a user-level threading library, which spawns a thread per request. We
use \texttt{httpaf} library~\cite{} for HTTP handling, and
\texttt{libev}~\cite{} for the eventloop. We compare our implementation against
an Lwt version (|lwt|) which also uses \texttt{httpaf} and \texttt{libev}.
Unlike using effect handlers, Lwt version is written in monadic style, and does
not have a notion of thread per request. For comparison, we also include a Go
version (|go|) that uses \texttt{net/http}~\cite{} package.

\begin{figure}
	\begin{minipage}{0.49\linewidth}
		\centering
		\includegraphics[width=\linewidth]{http-benchmarks/throughput.pdf}
		\subcaption{Throughput}
		\label{grf:throughput}
	\end{minipage}
	\begin{minipage}{0.49\linewidth}
		\includegraphics[width=\linewidth]{http-benchmarks/latency.pdf}
		\subcaption{Tail latency}
		\label{grf:latency}
	\end{minipage}
	\mycaption{Web server performance.}
	\vspace{-2mm}
\end{figure}

The client workload was generated with \texttt{wrk2}~\cite{}. We maintain 1k
open connections and perform requests for a static web page at different rates,
and record the request service rate and latency. The throughput and tail
latency graphs are given in Figures~\ref{grf:throughput} and~\ref{grf:latency}.
In all the versions, the throughput plateaus at around 30k requests per second.
We measure the tail latencies at 2/3rd of this rate (at 20k requests per
second), as would be the case in real deployments. We observe that both the
OCaml versions remain competitive with |go|, and |MC| performs best in terms of
tail latency.

Using effect handlers in the web server also helps with debugging the system as
a whole. Multicore OCaml supports getting the backtrace of a continuation,
which is a minor extension to getting the current backtrace. We can use this
feature to dump the backtrace of all the user-level threads in the web server.
Such a feature is provided by Go~\cite{}, but not possible with OCaml
concurrency libraries such as Async and Lwt which lack the notion of a thread.

@@
